### 2021-07-25 tf_prediction.tex
code implementation using tensorflow

\begin{minted}{python}
import tensorflow as tf
from keras_preprocessing.image import img_to_array
from tensorflow.keras.preprocessing.image import ImageDataGenerator
from tensorflow.keras.applications.resnet50 import preprocess_input, decode_predictions
from tensorflow.keras.preprocessing import image
from tensorflow.keras.preprocessing import image_dataset_from_directory
import matplotlib.image as mpimg
from keras.preprocessing import image
import cv2
import os
import glob

import numpy as np
import matplotlib.pyplot as plt

model = tf.keras.models.load_model('/content/gdrive/MyDrive/model/vgg_imp.h5')

training_dataset_directory = '/content/dataset/training'
testing_dataset_directory = '/content/dataset/testing'

train_datagen = ImageDataGenerator(
    rotation_range=20,
    width_shift_range=0.2,
    height_shift_range=0.2,
    horizontal_flip=True,
    vertical_flip=True,
    preprocessing_function=preprocess_input,
    validation_split=0.2
)

img_size = (150, 150)
img_shape = img_size + (3,)
random_seed = 123

training_set = train_datagen.flow_from_directory(
    training_dataset_directory,
    target_size=img_size,
    class_mode='binary',
    subset='training'
)

testing_set = train_datagen.flow_from_directory(
    testing_dataset_directory,
    target_size=img_size,
    class_mode='binary',
    subset='validation'
)

loss, acc = model.evaluate(training_set)

IMG_SIZE = (150,150)
BATCH_SIZE = 32
new_dataset = image_dataset_from_directory(
    '/content/gdrive/MyDrive/dataset',
    image_size=IMG_SIZE
)

predictions = model.predict(new_dataset,batch_size=BATCH_SIZE)

img_dir = "/content/gdrive/MyDrive/dataset/pred"
data_path = os.path.join(img_dir,'*g')
files = glob.glob(data_path)
data = []
result = []
for f1 in files:
    test_image = image.load_img(f1, target_size = (150, 150))
    test_image = image.img_to_array(test_image)
    test_image = np.expand_dims(test_image, axis = 0)
    images = np.vstack([test_image])
    classes = model.predict(images, batch_size=10)
    classes = np.round(classes)
    data.append(f1)
    result.append(classes)

for x,y in zip(data, result):
  print(x, y)

img = cv2.imread(data[0])
plt.imshow(img)
plt.title(result[0])

L = 4
W = 4
fig, axes = plt.subplots(L, W, figsize = (12,12))
axes = axes.ravel()
for i in np.arange(0, L * W):
    img = cv2.imread(data[i])
    axes[i].imshow(img)
    axes[i].set_title(result[i])
    axes[i].set_xticks([])
    axes[i].set_yticks([])
\end{minted}