### 2021-07-25 tf_vgg.tex

\begin{minted}{python}
import tensorflow as tf
from tensorflow.keras.preprocessing.image import ImageDataGenerator
from tensorflow.keras.applications.vgg16 import VGG16, preprocess_input
from tensorflow.keras.models import Sequential
from tensorflow.keras.layers import Flatten, Dropout, Dense
from matplotlib import pyplot as plt

training_dataset_directory = '/content/dataset/training'
testing_dataset_directory = '/content/dataset/testing'

train_datagen = ImageDataGenerator(
    rotation_range=20,
    width_shift_range=0.2,
    height_shift_range=0.2,
    horizontal_flip=True,
    vertical_flip=True,
    preprocessing_function=preprocess_input,
    validation_split=0.2
)

img_size = (150, 150)
img_shape = img_size + (3,)
random_seed = 123

training_set = train_datagen.flow_from_directory(
    training_dataset_directory,
    target_size=img_size,
    class_mode='binary',
    subset='training'
)

testing_set = train_datagen.flow_from_directory(
    testing_dataset_directory,
    target_size=img_size,
    class_mode='binary',
    subset='validation'
)

img_size = (150, 150)
img_shape = img_size + (3,)
random_seed = 123

vgg16 = VGG16(
    weights='imagenet',
    include_top=False,
    input_shape=img_shape
)

num_classes = 1
model = Sequential()
model.add(vgg16)
model.add(Flatten())
model.add(Dense(256, activation='relu'))
model.add(Dropout(0.5))
model.add(Dense(num_classes, activation='sigmoid'))
model.layers[0].trainable = False
model.summary()

model.compile(
    loss='binary_crossentropy',
    optimizer='Adam',
    metrics=['accuracy']
)

epochs = 30
history = model.fit(
    training_set,
    epochs=epochs,
    validation_data=testing_set,
)
\end{minted}