### 2021-07-25 fastai_resnet.tex
resnet and vgg using fastai

\begin{minted}{python}
import numpy as np
import pandas as pd
import torch
import keras
import os
import gc
import pathlib
from sklearn.metrics import confusion_matrix
from fastai import *
from fastai.vision.all import *
from fastai.tabular.all import *
from fastai.text.all import *
from fastai.vision.data import ImageDataLoaders
from fastai.vision.models import *
import torchvision.models as models
from fastai.callback.schedule import lr_find
from fastai.callback.schedule import *
from matplotlib import pyplot as plt
from fastai.imports import *
from fastai.torch_core import *
from fastai.learner import *

print(os.listdir("/content/drive/MyDrive/Colab Notebooks/samples"))
DATA_DIR = "/content/drive/MyDrive/Colab Notebooks/samples/Brain_Tumor"
os.listdir(f'{DATA_DIR}')

data = ImageDataLoaders.from_folder(
    DATA_DIR,
    train=".",
    valid_pct=0.2,
    ds_tfms=aug_transforms(
        mult=1.0,
        do_flip=True,
        flip_vert=True,
        max_warp=0,
        max_rotate=10.0,
        max_zoom=1.1,
        max_lighting=0.2,
        p_affine=0.75,
        mode='bilinear',
        pad_mode='reflection',
        align_corners=True,
        min_scale=1.0),
    item_tfms=Resize(224),
    bs=64,
    val_bs=None,
    num_workers=0,
    batch_tfms=Normalize.from_stats(
        *imagenet_stats))

data.show_batch(nrows=4, figsize=(10, 10))

learner = cnn_learner(
    data, models.resnet50, metrics=[
        accuracy, error_rate], cbs=[
            ShowGraphCallback()], model_dir="/tmp/model/")

learner.lr_find()

learner.model

learner.fit_one_cycle(6, lr_max=slice(8.32e-4), cbs=[ShowGraphCallback()])

learner.export(
    "/content/drive/MyDrive/Colab Notebooks/pretrained_model/resnet/btc2.pkl")

learner.unfreeze()

learner.lr_find()

learner.fit_one_cycle(
    6,
    lr_max=slice(5.754399353463668e-6),
    cbs=[
        ShowGraphCallback()])

learner.fit_one_cycle(4, lr_max=slice(2.7e-5), cbs=[ShowGraphCallback()])

learner.fit_one_cycle(4, lr_max=slice(2.9e-5), cbs=[ShowGraphCallback()])

learner.export(
    "/content/drive/MyDrive/Colab Notebooks/pretrained_model/resnet/final_resnet/btc_final2.pkl")

interp = ClassificationInterpretation.from_learner(learner)

interp.plot_top_losses(25, figsize=(20, 20))

interp.plot_confusion_matrix(figsize=(8, 8), dpi=60)
\end{minted}