### 2021-07-25 fastai_cnn_pred.tex

\begin{minted}{python}
import numpy as np
import pandas as pd
import fastbook
from fastbook import *
import os
import gc
import pathlib
import fastai.losses
import fastai.layers
from sklearn.metrics import confusion_matrix
from fastai import *
from fastai.vision import *
from fastai.vision.models import *
import torchvision.models as models

learner = load_learner(
    "/content/drive/MyDrive/Colab Notebooks/pretrained_model/cnn/btc4_final.pkl")

def is_true(x):
    if x[0] == 'yes':
        return "Tumor detected."
    else:
        return "Tumor not detected."

img = PILImage.create(
    f"/content/drive/MyDrive/Colab Notebooks/BrainTumor/no/2 no.jpeg")

x = learner.predict(img)

print("Test Result:", is_true(x), "\nProbability:", x[2])
\end{minted}