### 2021-07-31 acknowledgement.tex
changed first 3 pages as aksed

\section*{\hfill \Huge Acknowledgement \hfill}%
\label{sec:Acknowledgement}

% page border
\AddToShipoutPictureBG*{%
\begin{tikzpicture}[overlay,remember picture]
\draw[line width=1.5pt]
	($ (current page.north west) + (1cm,-1cm) $)
	rectangle
	($ (current page.south east) + (-1cm,2cm) $);
\end{tikzpicture}
}

\vspace{3cm}

\onehalfspacing

I take this opportunity to express profound sense of gratitude and respect to
all those who helped me throughout the duration of this project.

\vspace{1cm}

First and foremost, I would like to thank my supervisor of this project, Prof.
\textbf{Kajari Bhattacharjee}, for her valuable time and guidance. Her
willingness to motivate me has contributed tremendously to my project.

\vspace{1cm}

I would also like to thank Prof. \textbf{Srijeeta Chakrabarti}, Head of
Department of Computer Science, for providing with a good environment and
facilities to work on this project.

\vspace{1cm}
%\vfill

I regret any inadvertent omissions.

%\vspace{3cm}
\vfill

\rule{7cm}{0.4pt}

\textbf{Brahmajit Das} - 183313-21-0044

Semester VI (Honours)

Gurudas College

University of Calcutta

Kolkata - 700 054

\singlespacing