### 2021-07-31 acknowledgement.tex
text in italic

\section*{}%

\begin{center}
	{
		\Huge
		\textbf{Acknowledgement}
	}
\end{center}

% page border
\AddToShipoutPictureBG*{%
\begin{tikzpicture}[overlay,remember picture]
\draw[line width=1.5pt]
	($ (current page.north west) + (1cm,-1cm) $)
	rectangle
	($ (current page.south east) + (-1cm,2cm) $);
\end{tikzpicture}
}

\vspace{3cm}

\onehalfspacing

\textit{
	I take this opportunity to express profound sense of gratitude and respect to
	all those who helped me throughout the duration of this project.
}

\vspace{1cm}

\textit{
	First and foremost, I would like to thank my supervisor of this project, Prof.
	\textbf{Kajari Bhattacharjee}, for her valuable time and guidance. Her
	willingness to motivate me has contributed tremendously to my project.
}

\vspace{1cm}

\textit{
	I would also like to thank Prof. \textbf{Srijeeta Chakrabarti}, Head of
	Department of Computer Science, for providing with a good environment and
	facilities to work on this project.
}

\vspace{1cm}
%\vfill

\textit{
	I regret any inadvertent omissions.
}

%\vspace{3cm}
\vfill

\rule{7cm}{0.4pt}

\textbf{Brahmajit Das} - 183313-21-0044

Semester VI (Honours)

Gurudas College

University of Calcutta

Kolkata - 700 054

\singlespacing