### 2021-08-01 paper.tex
added a few subsections and subsubsections

\documentclass[10pt]{report}

\input{imports.tex}

\begin{document}

	\pagenumbering{roman}
	% Title page
	\input{titlepage.tex}

	% certificate page
	\input{certificate.tex}

	% acknowledgement
	\input{acknowledgement.tex}

	\tableofcontents

	% abstract
	%\section[abstract]{Abstract}
	\chapter{Abstract}

	% reverting back to normal page numbering
	\pagenumbering{arabic}

	\thispagestyle{plain}

	Tumors are cancerous or non-cancerous mass or growth of abnormal cells in
	brain. Tumors can start in brain, or cancer elsewhere in the body can spread
	to brain. There are many way to control the occurrence of these abnormal
	cells. A tumor can be denoted as a malformed mass of tissues wherein the
	cells multiply abruptly and ceaselessly, that is there is no control over
	the growth of the cells.

	The process of Image segmentation is adopted for extracting abnormal tumor
	region within the brain. In the MRI (magnetic resonance image), segmentation
	of brain tissue holds very significant in order to identify the presence of
	outlines concerning the brain tumor. There is abundance of hidden
	information in stored in the Health care sector. With appropriate use of
	accurate data mining classification techniques, early prediction of any
	disease can be effectively performed.

	The project examines list of risk factors that are being traced out in brain
	tumor surveillance systems. Also the method proposed assures to be highly
	efficient and precise for brain tumor detection, classification and
	segmentation. To achieve this precise automatic or semi-automatic methods
	are needed. The project proposes an automatic segmentation method that
	relies upon \textit{ CNN (Convolution Neural Networks) }, \textit{ VGG 16 }
	and \textit{ Resent 50 }, determining small 7 x 7 kernels. By incorporating
	this single technique, segmentation and classification is accomplished. CNN
	(a ML technique) from NN (Neural Networks)wherein it has layer based for
	results classification.

	Various levels involved in the proposed mechanisms are:

	\begin{enumerate}
		\item Data collection
		\item Pre-processing
		\item Average filtering
		\item Segmentation
		\item Feature extraction
		\item CNN (or any other model) via classification and identification. By
			utilizing the DM (data mining) techniques, significant relations and
			patterns from the data can be extracted. The techniques of ML
			(machine learning) and Data mining are being effectively employed
			for brain tumor detection and prevention at an early stage.
	\end{enumerate}

	\chapter{Introduction}
	\section[domain description]{Domain Description}

	\begin{enumerate}
		\item \textbf{Neurological Examination}: It is a series of test to
			measures the function of the patients nervous system and also
			his/her physical and mental alertness.
		\item \textbf{Machine Learning}: Machine learning approaches address
			these problems by mainly using hand-crafted features (or pre-defined
			features). As an initial step in this kind of segmentation, the key
			information is extracted from the input image using some feature
			extraction algorithm, and then a discriminative model is trained to
			recognize the tumor from normal tissues. The designed machine
			learning techniques generally employ hand-crafted features with
			various classifiers, such as random forest, support vector machine
			(SVM), fuzzy clustering. The designed methods and features
			extraction algorithms have to extract features, edge-related
			details, and other necessary information—which is time-consuming.
			Moreover, when boundaries between healthy tissues and tumors are
			fuzzy/vague, these methods demonstrate poorer performances.
		\item \textbf{Brain Scan}: Brain scan is a picture of the internal
			structure of the brain. A specialized machine takes a scan in the
			same way as a digital camera takes a photograph. Using computer
			technology, a scan compiles an image of the brain by photographing
			it from various angles. Some types of scan uses contrast agent (or
			contrast dye), which helps the doctor to see the difference between
			normal and abnormal brain tissues.

			MRI (Magnetic Resonance Imaging): It is a scanning device that uses
			magnetic field and computer to capture images of the brain on films.
			It does not use x-rays. It provides pictures from various planes,
			which permits doctor to create a three-dimensional image of the
			tumor. The MRI detects signals emitted from normal and abnormal
			tissues, providing clear images of almost all tumors.
	\end{enumerate}

	\section[motivation]{Motivation}
	The motivation is to develop a software with better segmentation capability
	for use in medical imaging to detect diseases like brain tumor. Image
	segmentation has been identified as the key problem of medical image
	analysis and remains a popular and challenging area of research. Image
	segmentation is increasingly used in many clinical and research applications
	to analyze medical imaging datasets; which motivated us to present a
	snapshot of dynamically changing field of medical image segmentation.

	CT (Computed Tomography), MRI (Magnetic Resonance Imaging), PET (Positron
	Emission Tomography) etc. generates a large amount of image information.
	With the improved technology, not only does the size and resolution of the
	images grow but also the number of dimensions increases. In the future, we
	would like to have algorithms which can automatically detect diseases,
	lesions and tumors, and highlight their locations in the large pile of
	images.

	The motivation of this work is to increase patient safety by providing
	better and more precise data for medical decision.

	\section[scope of work]{Scope of Work}
	\subsection[deliverables]{Deliverables}
	\begin{itemize}
		\item Working program to take an MRI scan as input and predict presence
			of tumorous cells with $\geq 90\%$ accuracy.
	\end{itemize}

	\section[scope]{Scope}
	\begin{itemize}
		\item The working program has external dependencies ( libraries ) and
			it's expected to have a them installed for the program to work.
	\end{itemize}

	\subsubsection[timeline]{Timeline}
	\begin{itemize}
		\item \textbf{April 27, 2021} Project Assigned
		\item \textbf{May 2, 2021} Project finalized by supervisor, and group is
			divided into groups of two.
		\item \textbf{May 3, 2021} Data collection started.
		\item \textbf{May 12, 2021} Project Repository created and coding is
			started.
		\item \textbf{July 7, 2021} Coding is finished, documentation is
			started.
		\item \textbf{Just 21, 2021} Documentation complete.
	\end{itemize}

	\subsubsection[reports]{Reports}
	\begin{itemize}
		\item Constantly updating and pushing code to repository.
		\item Both teams staying in touch with each other to keep up with each
			others progress.
		\item Report back to supervisor every once a week.
	\end{itemize}

	\chapter{Background}
	\input{backgournd.tex}

	\chapter{Methodology}

	As per literature survey, it was found that automated  brain  tumor
	detection  is  very  necessary  as  high accuracy is needed when human life
	is involved.  Automated detection of tumor in MR images involves feature
	extraction and  classification  using  machine  learning  algorithm.  In
	this paper, a system to automatically detect tumor in MR images is proposed
	as shown in Figure~\ref{Proposed Methodology}

	\begin{figure}[h]
		\centering
		\includegraphics[width=1\linewidth]{imgs/methodology}
		\caption{Proposed Methodology}%
		\label{Proposed Methodology}
	\end{figure}

	\section{Image Acquisition}%

	The MRI brain images are acquired and are
	given as input to pre-processing stage. The sample brain MR images are shown
	in Figure~\ref{MRIScans}.

	\begin{figure}[H]
		\centering
		\subfloat[MRI scan shown no presence of tumor]
		{\includegraphics[width=0.4\linewidth]
		{imgs/pred1.jpg}\label{scan2}}
		\hspace{0.5cm}
		\subfloat[MRI scan of a tumorous cell]
		{\includegraphics[width=0.3\linewidth]
		{imgs/Y100.JPG}\label{scan1}}
		\caption{MRI Scans}
		\label{MRIScans}
	\end{figure}

	\section{Preprocessing}%

	Preprocessing is needed as it provides improvement
	in image data which  enhances some of the image features which are important
	for further processing. The  pre-processing steps that are applied to MR
	image are as follows:

	\begin{enumerate}
		\item The  RGB  MR    image  is  converted  to  gray  scale image  and
			then  median  filter  is  applied  for  noise  removal from brain MR
			images as shown in Figure~\ref{smooted image}. The noise is to removed for further
			processing as high accuracy is needed.

		\item Then  edges  are  detected  from  filtered  image  using canny
			edge  detection  as  shown  in  Figure~\ref{edge detection}.  The  edge detected
			image is needed for segmentation of  the image

		\item Then  watershed  segmentation  is  done  for  finding the
			location  of  the  tumor  in  the  brain  image  as  shown  in
			Figure~\ref{segmentation}. Segmentation is  the process of dividing  an image into
			multiple segments. The aim of segmentation is to change
			representation  of    image  into  something  which  is more  easy
			to  analyze.  The result  of  watershed  segmentation is  label
			image.    In  label  image,  all  the  different  objects identified
			will  have  different  pixel  values  ,  all  the  pixels  of first
			object  will have value 1, all the pixels of second object will have
			value 2 and so on.
	\end{enumerate}

	\begin{figure}[ht]
		\centering
		\subfloat[Original Image]
		{\includegraphics[width=0.2\linewidth]
		{imgs/Y100.JPG}\label{original}}
		\hspace{0.5cm}
		\subfloat[Filtered Imaage]
		{\includegraphics[width=0.2\linewidth]
		{imgs/median_filtering.png}\label{smooted image}}
		\hspace{0.5cm}
		\subfloat[Edge Detection]
		{\includegraphics[width=0.2\linewidth]
		{imgs/edges_detection.png}\label{edge detection}}
		\hspace{0.5cm}
		\subfloat[Segmentation]
		{\includegraphics[width=0.2\linewidth]
		{imgs/segmentation.png}\label{segmentation}}
		\caption{preprocessing operations}
		\label{Preprocessing Operations}
	\end{figure}

	\section{Feature Extraction}%

	When input to an algorithm is very large and
	redundant to be processed, it is transformed into reduced representative set
	of features called feature vector. Transformation of input data into set of
	features is called feature extraction. In this step, the important features
	needed for image classification are extracted. The segmented brain MR image
	is used and texture features are extracted from the segmented image which
	shows the texture property of the image. These features are extracted using
	Gray Level Co-occurrence Matrix (GLCM) as it is robust method with high
	performance.

	The GLCM features are extracted as follows:

	\begin{enumerate}
		\item Energy: It gives a measure of textural uniformity, that is, measure
			of pixel pair repetitions.

			\begin{equation}
				E = \sum^{N_g - 1}_{i=0} \sum^{N_g - 1}_{j=0} p(i, j)^2 \
				\textnormal{ here, range = [0, 1] }
			\end{equation}

		\item Contrast: It gives a measure of intensity contrast between a pixel
			and its neighbor over the whole image.

			\begin{equation}
				Con = \sum^{N_g -1}_{n=0} n^2 \sum^{N_g -1}_{i=0} \sum^{N_g
				-1}_{j=0} p(i,j)^2 \
				\textnormal{ here, range = [0, 1] }
			\end{equation}

		\item Correlation: It gives a measure of how correlated a pixel to its
			neighbor over the whole image.

			\begin{equation}
				C = \frac{1}{\sigma^x \sigma^y}
				\sum^{N_g - 1}_{i=0} \sum^{N_g -1}_{j=0} (i,j) p(i,j)^2
				- \mu_x \mu_y \ \textnormal{here, range = [-1, 1]}
			\end{equation}

		\item Homogeneity: It gives a measure of closeness of distribution of
			elements in GLCM to GLCM diagonal.

			\begin{equation}
				H = \sum^{N_g -1}_{i=0} \sum^{N_g -1}_{j=0}
				\frac{p(i,j)}{1 + \pmod{i,j}}
				\ \textnormal{here, range = [0, 1]}
			\end{equation}
	\end{enumerate}

	\section{Classification}%

	The Machine learning algorithms are used for
	classification of MR brain image either as normal or abnormal.  The major
	aim of ML algorithms is to automatically learn and make intelligent
	decisions. The feature set formed by above specified method was applied to
	Multi-Layer Perceptron (MLP) and Naive Bayes for classification.  MLP is
	a feed forward artificial neural network model that maps sets of input data
	into a set of appropriate output. It is known as feed forward because it
	does not contain any cycles and network output depends only on the current
	input instance. In MLP, each node is a neuron with a nonlinear activation
	function. It is based on supervised learning technique. Learning take place
	by changing connection weights after each piece of data is processed, based
	on the amount of error in the target output as compared to the expected
	result. The goal of the learning procedure is to minimize error by improving
	the current values of the weight associated with each edge. Because of this
	backward changing process of the weights, model is named as
	back-propagation.

	Naive bayes is a supervised learning as well as statistical method for
	classification. It is simple probabilistic classifier based on Bayes
	theorem. It assumes that the value of a particular feature is unrelated to
	the presence or absence of any other feature. The prior probability and
	likelihood are calculated in order to calculate the posterior probability.
	The method of maximum posterior probability is used for parameter
	estimation. This method requires only a small amount of training data to
	estimate the parameters which are needed for classification. The time taken
	for training and classification is less.

	For classification three models are used; \textbf{CNN}, \textbf{VGG 16} and
	\textbf{Resnet 50}.

	\section{Convolutional Neural Network}%

	CNN or Convolutional Neural Network a class of artificial neural network,
	most commonly applied to analyze visual imagery.They are also known as
	shift invariant or space invariant artificial neural networks (SIANN), based
	on the shared-weight architecture of the convolution kernels or filters that
	slide along input features and provide translation equivariant responses
	known as feature maps. Counter-intuitively, most convolutional neural
	networks are only equivariant, as opposed to invariant, to translation.
	They have applications in image and video recognition, recommender
	systems, image classification, image segmentation, medical image
	analysis, natural language processing, brain-computer interfaces, and
	financial time series.

	\begin{figure}[h]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/cnn_arch.png}
		\caption{CNN architecture}%
		\label{CNN Architecture}
	\end{figure}

	\subsection{Architecture}

	A convolutional neural network consists of an input layer, hidden layers and
	an output layer. In any feed-forward neural network, any middle layers are
	called hidden because their inputs and outputs are masked by the activation
	function and final convolution. In a convolutional neural network, the
	hidden layers include layers that perform convolutions. Typically this
	includes a layer that performs a dot product of the convolution kernel with
	the layer's input matrix.  This product is usually the Frobenius inner
	product, and its activation function is commonly ReLU. As the convolution
	kernel slides along the input matrix for the layer, the convolution
	operation generates a feature map, which in turn contributes to the input of
	the next layer. This is followed by other layers such as pooling layers,
	fully connected layers, and normalization layers \cite{covnet_arch}.

	\subsection{Convolutional Layer}:

	In a CNN, the input is a tensor with a shape: (number of inputs) x (input
	height) x (input width) x (input channels). After passing through a
	convolutional layer, the image becomes abstracted to a feature map, also
	called an activation map, with shape: (number of inputs) x (feature map
	height) x (feature map width) x (feature map channels). A convolutional
	layer within a CNN generally has the following attributes:

	\begin{itemize}
		\item Convolutional filters/kernels defined by a width and height (hyper-parameters).
		\item The number of input channels and output channels
			(hyper-parameters). One layer's input channels must equal the number
			of output channels (also called depth) of its input.
		\item Additional hyperparameters of the convolution operation, such as:
			padding, stride, and dilation.
	\end{itemize}

	Convolutional layers convolve the input and pass its result to the next
	layer. This is similar to the response of a neuron in the visual cortex to a
	specific stimulus.

	\subsection{Pooling Layer}:

	Convolutional networks may include local and/or global pooling layers along
	with traditional convolutional layers. Pooling layers reduce the dimensions
	of data by combining the outputs of neuron clusters at one layer into a
	single neuron in the next layer. Local pooling combines small clusters,
	tiling sizes such as 2 x 2 are commonly used. Global pooling acts on all the
	neurons of the feature map. There are two common types of pooling in popular
	use: max and average. Max pooling uses the maximum value of each local
	cluster of neurons in the feature map, while average pooling takes the
	average value.

	\subsection{ReLU Layer}:

	ReLU is the abbreviation of rectified linear unit, which applies the
	non-saturating activation function
	\begin{equation}
		f ( x ) = \max ( 0 , x )
	\end{equation}

	.It effectively removes
	negative values from an activation map by setting them to zero. It
	introduces nonlinearities to the decision function and in the overall
	network without affecting the receptive fields of the convolution layers.

	Other functions can also be used to increase nonlinearity, for example the
	saturating hyperbolic tangent
	\begin{equation}
	f(x)=\tanh(x)
	f(x) = | \tanh (x) |
	\end{equation}
	and the
	sigmoid function

	\begin{equation}
		\sigma(x) = (1+e^{-x})^{-1}
	\end{equation}

	ReLU is often preferred to other functions because it trains the neural network several
	times faster without a significant penalty to generalization accuracy
	 \cite{relu_layer}.

	\subsection{Fully Connected Layer}

	After several convolutional and max pooling layers, the final classification
	is done via fully connected layers. Neurons in a fully connected layer have
	connections to all activations in the previous layer, as seen in regular
	(non-convolutional) artificial neural networks. Their activations can thus
	be computed as an affine transformation, with matrix multiplication followed
	by a bias offset (vector addition of a learned or fixed bias term).

	\subsection{Loss Layer}

	The "loss layer", or "loss function", specifies how training penalizes the
	deviation between the predicted output of the network, and the true data
	labels (during supervised learning). Various loss functions can be used,
	depending on the specific task.

	The Softmax loss function is used for predicting a single class of K
	mutually exclusive classes. Sigmoid cross-entropy loss is used for
	predicting K independent probability values is $[0, 1]$.  Euclidean loss is
	used for regressing to real-valued labels $(- \infty, \infty)$.

	\subsection{Receptive field}

	In neural networks, each neuron receives input from some number of locations
	in the previous layer. In a convolutional layer, each neuron receives input
	from only a restricted area of the previous layer called the neuron's
	receptive field. Typically the area is a square (e.g. 5 by 5 neurons).
	Whereas, in a fully connected layer, the receptive field is the entire
	previous layer. Thus, in each convolutional layer, each neuron takes input
	from a larger area in the input than previous layers. This is due to
	applying the convolution over and over, which takes into account the value
	of a pixel, as well as its surrounding pixels. When using dilated layers,
	the number of pixels in the receptive field remains constant, but the field
	is more sparsely populated as its dimensions grow when combining the effect
	of several layers.

	\subsection{Weights}

	Each neuron in a neural network computes an output value by applying a
	specific function to the input values received from the receptive field in
	the previous layer. The function that is applied to the input values is
	determined by a vector of weights and a bias (typically real numbers).
	Learning consists of iteratively adjusting these biases and weights.

	The vector of weights and the bias are called filters and represent
	particular features of the input (e.g., a particular shape). A
	distinguishing feature of CNNs is that many neurons can share the same
	filter. This reduces the memory footprint because a single bias and a single
	vector of weights are used across all receptive fields that share that
	filter, as opposed to each receptive field having its own bias and vector
	weighting.

	\section{VGG 16}

	VGG 16 is is a significantly more accurate ConvNet architecture, which not
	only achieve state-of-the-art accuracy on ILSVRC classification and
	localisation tasks, but are also applicable to other image recognition
	datasets, where they achieve excellent performance even when used as a part
	of a relatively simple pipelines (e.g. deep features classified by a linear
	SVM without fine-tuning) \cite[1]{simonyan2015deep}.

	\subsection{Architecture}
	\label{sec:vgg_arch}

	\begin{wrapfigure}{r}{0.5\textwidth}
		\begin{center}
			\includegraphics[ width=0.5\textwidth ]{imgs/vgg16_arch.png}
		\end{center}
		\caption{VGG Architecture}
	\end{wrapfigure}

	During training, the input to our ConvNets is a fixed-size 224 × 224 RGB
	image. The only pre- processing we do is subtracting the mean RGB value,
	computed on the training set, from each pixel.  The image is passed through
	a stack of convolutional (conv.) layers, where we use filters with a very
	small receptive field: 3 × 3 (which is the smallest size to capture the
	notion of left/right, up/down, center). In one of the configurations we also
	utilise 1 × 1 convolution filters, which can be seen as a linear
	transformation of the input channels (followed by non-linearity). The
	convolution stride is fixed to 1 pixel; the spatial padding of conv. layer
	input is such that the spatial resolution is preserved after convolution,
	i.e. the padding is 1 pixel for 3 × 3 conv. layers. Spatial pooling is
	carried out by five max-pooling layers, which follow some of the conv.
	layers (not all the conv. layers are followed by max-pooling). Max-pooling
	is performed over a 2 × 2 pixel window, with stride 2.

	A stack of convolutional layers (which has a different depth in different
	architectures) is followed by three Fully-Connected (FC) layers: the first
	two have 4096 channels each, the third performs 1000- way ILSVRC
	classification and thus contains 1000 channels (one for each class). The
	final layer is the soft-max layer. The configuration of the fully connected
	layers is the same in all networks.

	All hidden layers are equipped with the rectification (ReLU) non-linearity.
	We note that none of our networks (except for one) contain Local Response
	Normalisation (LRN) normalisation: as will be shown in Sect. 4, such
	normalisation does not improve the performance on the ILSVRC dataset, but
	leads to increased memory con- sumption and computation time.
	\cite[1]{simonyan2015deep}

	\subsection{Configuration}

	The ConvNet configurations, evaluated in this paper, are outlined in
	Figure~\ref{fig:vgg16_conf}, one per column. In the following we will refer
	to the nets by their names (A–E).  All configurations follow the generic
	design presented in Section~\ref{sec:vgg_arch} and differ only in the
	depth: from 11 weight layers in the network A (8 conv. and 3 FC layers) to
	19 weight layers in the network E (16 conv. and 3 FC layers). The width of
	conv. layers (the number of channels) is rather small, starting from 64 in
	the first layer and then increasing by a factor of 2 after each max-pooling
	layer, until it reaches 512.
	\cite[3]{simonyan2015deep}

	\begin{figure}[ht]
		\centering
		\includegraphics[width=0.5\linewidth]{imgs/vgg16_conf}
		\caption{VGG16 configuration}%
		\label{fig:vgg16_conf}
	\end{figure}

	\section{ResNet 50}%

	\subsection{Residual Learning}

	Let us consider $H(x)$ as an underlying mapping to be fit by a few stacked
	layers (not necessarily the entire net), with $x$ denoting the inputs to the
	first of these layers. If one hypothesizes that multiple nonlinear layers
	can asymptotically approximate complicated functions, then it is equivalent
	to hypothesize that they can asymptotically approximate the residual
	functions, i.e., $H(x) - x$ (assuming that the input and output are of the
	same dimensions). So rather than expect stacked layers to approximate $H(x)$
	, we explicitly let these layers approximate a residual function
	\begin{equation}
		F(x) = H(x) - x
	\end{equation}
	The original function thus becomes $F(x) + x$. Although both forms should be
	able to asymptot- ically approximate the desired functions (as
	hypothesized), the ease of learning might be different \cite[3]{he2015deep}.

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/resnet_arch.png}
		\caption{resnet50 configuration}%
		\label{fig:resnet conf}
	\end{figure}

	\subsection{Network Architecture}

	\begin{wrapfigure}{r}{0.5\textwidth}
		\begin{center}
			\includegraphics[ width=0.45\textwidth ]{imgs/resnet_arch2.png}
		\end{center}
		\caption{Resnet Architecture}
	\end{wrapfigure}

	The plain/residual network was tested as follows:

	\subsubsection{Plain Network}

	Our plain baselines are mainly inspired by the philosophy of VGG nets.  The
	convolutional layers mostly have 3×3 filters and follow two simple design
	rules:

	\begin{enumerate}
		\item for the same output feature map size, the layers have the same
			number of filters
		\item if the feature map size is halved, the number of filters is
			doubled so as to preserve the time complexity per layer.
	\end{enumerate}

	We perform downsampling directly by convolutional layers that have a stride
	of 2. The network ends with a global average pooling layer and a 1000-way
	fully-connected layer with softmax. The total number of weighted layers are
	34.  It is worth noticing that our model has fewer filters and lower
	complexity than VGG nets. Our 34-layer baseline has 3.6 billion FLOPs
	(multiply-adds), which is only 18\% of VGG-19 (19.6 billion FLOPs)
	\cite[3]{he2015deep} .

	\subsubsection{Residual Network}

	Based on the above plain network, we insert shortcut connections which turn
	the network into its counterpart residual version. The identity shortcuts
	can be directly used when the input and output are of the same dimensions.
	When the dimensions increase, we consider two options:
	\begin{enumerate}
		\item The shortcut still performs identity mapping, with extra zero
			entries padded for increasing dimensions. This option introduces no
			extra parameter
		\item The projection shortcut is used to match dimensions (done by 1×1
			convolutions).
	\end{enumerate}
	For both options, when the shortcuts go across feature maps of two sizes,
	they are performed with a stride of 2 \cite[4]{he2015deep} .

	\chapter{Implementation}

	\section{Assumptions and Dependencies}%
	It is assumed that the MRI scans are collected and processed before feeding
	into the module.

	The program is dependent on external modules and are expected to be
	pre-installed on the system. The modules namely include:

	\begin{itemize}
		\item \href{https://www.fast.ai/}{fast.ai}
		\item \href{https://www.tensorflow.org/}{tensorflow}
		\item \href{https://opencv.org/}{OpenCV}
		\item \href{https://matplotlib.org/}{matplotlib}
	\end{itemize}

	\section{Implementation Methods}%
	After acquiring the data, the data is classified into two classes (binary
	class) of yes, consisting of tumorous cell and no, not consisting tumorous
	cells. A second batch of images is kept separate for prediction, the
	prediction batch.

	Next three machine learning models are created. A \textbf{CNN} or a
	convolutional neural network, \textbf{VGG 16} and \textbf{Resnet 50}. The
	models are implemented in both fast.ai and tensorflow, to be compared later.

	While training the images are resized to $(150 \times 150)$ and augmented at
	runtime, thus removing bias as much as possible. Since there are only two
	possible classes (yes and no), a binary class is sufficient; dividing the
	images into training and testing (validation set) in $80:20$ ratio. Giving
	us $2603 \ \textnormal{and} \ 650$ images for the \textit{yes} and
	\textit{no} classes, respectively.

	After training is complete the models are then evaluated (further discussed
	in Chapter~\ref{chap:Res} )

	\chapter{Result}
	\label{chap:Res}

	The evaluated models are plotted on graph for better understanding of data.

	\begin{figure}[H]
		\centering
		\subfloat[cnn using tensorflow]
		{\includegraphics[width=0.4\linewidth]
		{imgs/tf_cnn.png}\label{tf_cnn_cmp}}
		\hspace{0.5cm}
		\subfloat[cnn using fast.ai]
		{\includegraphics[width=0.5\linewidth]
		{imgs/fastai_cnn.png}\label{fastai_vgg16_cmp}}
		\caption{CNN implemented in fast.ai and tensorflow}
		\label{CNN in fast.ai and tensorflow}
	\end{figure}

	\begin{figure}[H]
		\centering
		\subfloat[vgg16 using tensorflow]
		{\includegraphics[width=0.4\linewidth]
		{imgs/tf_vgg16.png}\label{tf_vgg16}}
		\hspace{0.5cm}
		\subfloat[vgg16 using fast.ai]
		{\includegraphics[width=0.5\linewidth]
		{imgs/fastai_vgg16.png}\label{fastai_vgg16}}
		\caption{VGG16 implemented in fast.ai and tensorflow}
		\label{VGG16 in fast.ai and tensorflow}
	\end{figure}

	\begin{figure}[H]
		\centering
		\subfloat[resnet50 using tensorflow]
		{\includegraphics[width=0.4\linewidth]
		{imgs/tf_resnet.png}\label{tf_resnet}}
		\hspace{0.5cm}
		\subfloat[resnet50 using fast.ai]
		{\includegraphics[width=0.5\linewidth]
		{imgs/fastai_resnet.png}\label{fastai_resnet}}
		\caption{Resnet 50 implemented in fast.ai and tensorflow}
		\label{ResNet 50 in fast.ai and tensorflow}
	\end{figure}

	We have achieved around 96\% and 98\% accuracy for for the VGG16 and ResNet,
	respectively; using \textit{tensorflow}, and 97\% (98\% after unfreezing)
	and 97\% (99\% after unfreezing) using \textit{fast.ai}. While using CNN
	accuracy was almost same when implemented by the both the libraries.

	There are lot of room for improvement but due to time and hardware
	constrains we were able to only implement till such.

	\begin{figure}[H]
		\centering
		\subfloat[prediction made by VGG16 implemented in tensorflow]
		{\includegraphics[width=0.5\linewidth]
		{imgs/result.png}\label{vgg16_tf}}
		\hspace{0.3cm}
		\subfloat[prediction made by ResNet50 implemented in fast.ai]
		{\includegraphics[width=0.5\linewidth]
		{imgs/result_fast_ai.png}\label{resnet_fastai}}
		\caption{Comparing Predictions made by models}
		\label{fig:Results_Comparison}
	\end{figure}

	As it can be seen from Figure~\ref{fig:Results_Comparison} there are some false
	positives in our predictions.

	\chapter{Appendix}

	\section{Tensorflow Implementation}%

	\section{CNN}
	\input{tf_cnn.tex}

	\section{VGG 16}
	\input{tf_vgg.tex}

	\section{ResNet 50}
	\input{tf_reset.tex}

	\section{Prediction}
	\input{tf_prediction.tex}

	\section{fast.ai Implementation}%

	\section{CNN}
	\input{fastai_cnn.tex}

	\textbf{Prediction}
	\input{fastai_cnn_pred.tex}

	\section{VGG 16}
	\input{fastai_vgg.tex}

	\textbf{Prediction}
	\input{fastai_vgg_pred.tex}

	\section{ResNet 50}
	\input{fastai_resnet.tex}

	\textbf{Prediction}
	\input{fastai_resnet_pred.tex}

	\emergencystretch=1em
	\printbibliography[heading=bibintoc,title={References}]

\end{document}