### 2021-08-03 fastai_cnn.tex

\begin{minted}{python}
import numpy as np
import pandas as pd
import torch
import keras
import tensorflow as tf
import os
import gc
import pathlib
from keras.models import Sequential
from keras.layers import Conv2D, MaxPooling2D, Flatten, Dense, Dropout, BatchNormalization
from sklearn.metrics import confusion_matrix
from fastai import *
from fastai.vision.all import *
from fastai.tabular.all import *
from fastai.text.all import *
from fastai.vision.data import ImageDataLoaders
from fastai.vision.models import *
import torchvision.models as models
from fastai.callback.schedule import lr_find
from fastai.callback.schedule import *
from matplotlib import pyplot as plt
from fastai.imports import *
from fastai.torch_core import *
from fastai.learner import *
from sklearn.model_selection import train_test_split
from sklearn.preprocessing import OneHotEncoder

print(os.listdir("/content/drive/MyDrive/Colab Notebooks/samples"))
DATA_DIR = "/content/drive/MyDrive/Colab Notebooks/samples/Brain_Tumor"
os.listdir(f'{DATA_DIR}')

data = ImageDataLoaders.from_folder(
    DATA_DIR,
    train=".",
    valid_pct=0.2,
    ds_tfms=aug_transforms(
        mult=2.0,
        do_flip=True,
        flip_vert=True),
    item_tfms=Resize(224),
    bs=64,
    batch_tfms=Normalize.from_stats(
        *imagenet_stats))

data.show_batch(nrows=4, figsize=(10, 10))

model = nn.Sequential(
    nn.Sequential(
        nn.Sequential(
            nn.Conv2d(3, 64, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.Conv2d(64, 64, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.MaxPool2d(2, stride=2),
            nn.Conv2d(64, 128, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.Conv2d(128, 128, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.MaxPool2d(2, stride=2),
            nn.Conv2d(128, 256, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.Conv2d(256, 256, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.MaxPool2d(2, stride=2),
            nn.Conv2d(256, 512, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.Conv2d(512, 512, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.Conv2d(512, 512, kernel_size=(3, 3),
			stride=(1, 1), padding=(1, 1)), nn.ReLU(inplace=True),
            nn.MaxPool2d(2, stride=2),
        )
    ),
    nn.Sequential(
        AdaptiveConcatPool2d(1),
        nn.Flatten(),
        nn.BatchNorm1d(1024, eps=1e-05, momentum=0.1,
		affine=True, track_running_stats=True),
        nn.Dropout(p=0.25, inplace=False),
        nn.Linear(in_features=1024, out_features=512, bias=False),
        nn.ReLU(),
        nn.BatchNorm1d(512, eps=1e-05, momentum=0.1,
		affine=True, track_running_stats=True),
        nn.Dropout(p=0.5, inplace=False),
        nn.Linear(in_features=512, out_features=2, bias=False)
    )
)

learner = Learner(data, model, metrics=[accuracy, error_rate])

learner.summary()

learner.lr_find()

learner.fit_one_cycle(13, lr_max=slice(1e-3), cbs=[ShowGraphCallback()])

learner.export(
    "/content/drive/MyDrive/Colab Notebooks/pretrained_model/cnn/btc4_final.pkl")

interp = ClassificationInterpretation.from_learner(learner)

interp.plot_top_losses(25, figsize=(20, 20))

interp.plot_confusion_matrix(figsize=(8, 8), dpi=60)
\end{minted}